---
title: "Morris Warmaking"
output:
  pdf_document: default
  df_print: paged
  html_document:
  word_document: default
author: Bernhard Koschicek
---
# War-making capacity 

## War-making capacity plot

Here out first dataset will be plotted and visualized. The y axis of the plot diagram is resized to 0.3 because of the explotion of warmaking capacity in the last 200 years. To better visualise the sublte changes across the centuries this was a necessary step. 

```{r} 
# Load csv
morris_warmaking <- read.delim("C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/morris_warmaking.csv")

# Asign variables
east = morris_warmaking$East
west = morris_warmaking$West
date = morris_warmaking$Date

#PDF output
#pdf(file="C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/figure.pdf", height=3.5, width=5)

# Create Plot
plot(date, east, type="o", col="brown", axes=FALSE, ann=FALSE, ylim=c(0,0.3))

# Name and config axis
axis(1, at=date, lab=date)
axis(2, at=east, las=1, tck=-0.05, cex.axis=0.7)
# Putting a box around 
box()

# Add data west
lines(date, west, type="o", pch=22, lty=6, col="violet")

# Labeling of the axis and diagram
title(main="War Data", col.main="green", font.main=3)
title(xlab="Date", col.lab=rgb(0,0.5,0))
title(ylab="War Readiness", col.lab=rgb(0,0.5,0))

# Creating the legend
legend(-4000, y= 0.3, c("east","west"), cex=0.8, col=c("brown","violet"), pch=21:22, lty=1:2)

#PDF output
#dev.off()
```

## Who was in lead 

This table represents who was in lead at the given time. We have a Date columne and tree columns for West, East and Equal. Whoever was in lead will get a True value.

```{r}
# Load csv
morris_warmaking <- read.delim("C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/morris_warmaking.csv")

# Asign variables
east = morris_warmaking$East
west = morris_warmaking$West
date = morris_warmaking$Date

# Make comparision and asign them to the values
west = morris_warmaking$West >  morris_warmaking$East
east = morris_warmaking$West <  morris_warmaking$East
equal = morris_warmaking$West ==  morris_warmaking$East

# Determine who was "better". I squeezed my brain for a better and fancy solution, but hours of thinking were pointless, so here is the simply not beautyful solution, I'm sorry. 

x = data.frame("Date" = morris_warmaking$Date, "West_better" = west, "East_better" = east,  "Equal" = equal)
x



```

# Maximum Settlement Sizes 

## Fix the dataset

> Fix the dataset so that it confirms to the principles of tidy data. What should be corrected?

Here are some points how I fixed my dataset. I'm sure I forgort things but well there is so much to do and so little time: 

* Get the whole thing in a table format with RegEx: `^(\W?\d+)\W(\S+)\s+([^\d]+)([^\s]+)\s+(.+)`
* The comma from the size column is interprated as decimal, so it has to be changed to . or nothing. 
* kill the author because he wrote "perhaps 500" in the western table for Mareybet and didn't asign a value for the points. In this case I just added the value NA to the table by myself and wrote the "perhaps" manually to the size, because I didn't want to change my regex just for a single entry
* One problem could be the 4 date of the western table because there the author starts with 8000 BCE. I would just ignore this and hope that the plot will to the right thing here and just start the eastern values at 4000 BCE 
* Commas are not supported in plot, therefore we have to change it and delete them
* Sometime there are inconsistencies like in the settlements column one time the author wrote ´London and Constantinople´ 

>The datasets for the East and the West are separate. What would be your strategies to combine them?

There are as always more ways to do this. I would clearly add another column with the header alignment (?!) to it and fill the cells with the values eastern/western, so I can use them in a single csv file. But out of convinience I just would import the files separetly for the visualisation. If we go really fancy and we don't have the alignment of the Settlements, I'm sure we could reference the Settlementnames with a gazetteer to get their geographical position and meta information. 


```{r} 
# Loading dataset and asign to variable

eastern <- read.csv("C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/settlements/morris_settlement_eastern.csv", sep=";")
western <- read.csv("C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/settlements/morris_settlement_western.csv", sep=";")
dates = western$Date

west_set = western$Settlement
west_size = western$Size
west_point = western$Points

east_set = eastern$Settlement
east_size = eastern$Size
east_point = eastern$Points

# Plot the population size (note the y axis is limited to 1,000,000 because of the population growth)
plot(dates, east_size, type="o", col="brown", axes=FALSE, ann=FALSE, ylim=c(0,1000000))

## Axis and boxing
axis(1, at=dates, lab=dates)
axis(2, at=east_size, las=1, tck=-0.05, cex.axis=0.7)
box()

## Add new data line
lines(dates, west_size, type="o", pch=22, col="violet")

## labeling
title(main="Population Size", col.main="black", font.main=3)
title(xlab="Date", col.lab=rgb(0,0.5,0))
title(ylab="Size", col.lab=rgb(0,0.5,0))

## Add legend
legend(-8000, y= 1000000, c("east","west"), cex=0.8, col=c("brown","violet"), pch=21:22, lty=1:2)


# Plot for the warmaking points
plot(dates, east_point, type="o", col="brown", axes=FALSE, ann=FALSE, ylim=c(0,10))

## Axis and boxing
axis(1, at=dates, lab=dates)
axis(2, at=east_point, las=1, tck=-0.05, cex.axis=0.7)
box()

## Add data line
lines(dates, west_point, type="o", pch=22, col="violet")

## labeling
title(main="Warmaking Points", col.main="black", font.main=3)
title(xlab="Date", col.lab=rgb(0,0.5,0))
title(ylab="Points", col.lab=rgb(0,0.5,0))

## Add legend
legend(-8000, y= 10, c("east","west"), cex=0.8, col=c("brown","violet"), pch=21:22, lty=1:2)


# Who was better and which Settlements were the top

## Comparison
west = west_point >  east_point
east = west_point <  east_point
equal = west_point == east_point

## Make a data frame
better_whole = data.frame("dates_lead" = dates, "West_better" = west, "East_better" = east,  "Equal" = equal, "Western_Cities" = west_set, "Eastern_Cities" = east_set)

# I think this is not a nice solution but I cannot make a valid if statement in a for each with data frames... this is frustrating... 

## Create dataset of whom was better at a given time 
western_data = data.frame("dates_lead" = dates, "West_better" = west, "Western_Cities" = west_set)
west_lead = subset(western_data,  West_better == TRUE)
west_lead$West_better = "W"
eastern_data = data.frame("dates_lead" = dates, "East_better" = east, "Eastern_Cities" = east_set)
east_lead = subset(eastern_data,  East_better == TRUE)
east_lead$East_better = "E"

d = merge(western_data, eastern_data, by.x = "dates_lead", sort = TRUE)
d

## Loading plyr package
library("plyr")
library("tidyr")
library("dplyr")

## Bringing the two tables together
total_lead = rbind.fill(west_lead, east_lead)
attach(total_lead)
newdata = total_lead[order(dates_lead),]
detach(total_lead)
# newdata
newdata[is.na(newdata)] <- ""

## Well I looked at the reading we had to tidy my tables and replace the NAs with an empty string to `unite` the datasets, since unite strangely doesn't kill NAs. So I found in the reading the replace_na function, sounded nice, but even the test code at the r documentation worked! ( https://www.rdocumentation.org/packages/tidyr/versions/0.8.3/topics/replace_na ) So I have no idea how to change NAs in fctr to an empty string. I searched a lot and tried a lot, but nothing changed really. At least some fishy code `newdata[is.na(newdata)] <- ""` worked on the chr data. 

# newdata %>% replace_na(list(Western_Cities = 0))
# newdata

uniteddata = unite(newdata,"Better", "West_better", "East_better", sep = "", remove = TRUE)
# u = unite(uniteddata, "Cities", "Western_Cities", "Eastern_Cities", sep = "", remove = TRUE)
# # u[is.na(u)] <- ""
uniteddata


```

# Datasets

```{r}
morris_warmaking
eastern
western
```