---
title: ""
output:
  html_document:Experiment with War Data
    df_print: paged
author: Bernhard Koschicek
---
# War-making capacity 

```{r} 
#Load csv
morris_warmaking <- read.delim("C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/morris_warmaking.csv")
#Variables
east = morris_warmaking$East
west = morris_warmaking$West
date = morris_warmaking$Date
#PDF output
#pdf(file="C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/figure.pdf", height=3.5, width=5)
#Create Plot
plot(date, east, type="o", col="brown", axes=FALSE, ann=FALSE, ylim=c(0,0.3))
#Name and config axis
axis(1, at=date, lab=date)
axis(2, at=east, las=1, tck=-0.05, cex.axis=0.7)
box()
#Add line
lines(date, west, type="o", pch=22, lty=6, col="violet")
#labeling
title(main="War Data", col.main="green", font.main=3)
title(xlab="Date", col.lab=rgb(0,0.5,0))
title(ylab="War Readiness", col.lab=rgb(0,0.5,0))
#Legend
legend(-4000, y= 0.3, c("east","west"), cex=0.8, col=c("brown","violet"), pch=21:22, lty=1:2)
#PDF output
#dev.off()
```

# Maximum Settlement Sizes 

> Fix the dataset so that it confirms to the principles of tidy data. What should be corrected?

* Get the whole thing in a table format with RegEx: ^(\W?\d+)\W(\S+)\s+([^\d]+)([^\s]+)\s+(.+)
* The comma from the size column is interprated as decimal, so it has to be changed to . or nothing. 
* kill the author because he wrote "perhaps 500" in the western table for Mareybet and didn't asign a value for the points. In this case I just added the value NA to the table by myself and wrote the "perhaps" manually to the size, because I didn't want to change my regex just for a single entry
* One problem could be the 4 date of the western table because there the author starts with 8000 BCE. I would just ignore this and hope that the plot will to the right thing here and just start the eastern values at 4000 BCE 

>The datasets for the East and the West are separate. What would be your strategies to combine them?

There are as always more ways to do this. I would clearly add another column with the header alignment (?!) to it and fill the cells with the values eastern/western, so I can you them in a single csv file. But out of convinience I just would import the files separetly for the visualisation. If we go really fancy and we don't have the alignment of the Settlements, I'm sure we could reference the Settlementnames with a gazetteer to get their geographical position and meta information. 


```{r} 
eastern <- read.csv2("C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/morris_settlement.csv")
head(morris_settlement)
western <- read.csv("C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/morris_settlement_wester.csv", sep=";")
head(morris_settlement_wester)

```