---
title: ""
output:
  pdf_document: default
  df_print: paged
  html_document:
  word_document: default
author: Bernhard Koschicek
---
# War-making capacity 

```{r} 
#Load csv
morris_warmaking <- read.delim("C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/morris_warmaking.csv")
#Variables
east = morris_warmaking$East
west = morris_warmaking$West
date = morris_warmaking$Date
#PDF output
#pdf(file="C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/figure.pdf", height=3.5, width=5)
#Create Plot
plot(date, east, type="o", col="brown", axes=FALSE, ann=FALSE, ylim=c(0,0.3))
#Name and config axis
axis(1, at=date, lab=date)
axis(2, at=east, las=1, tck=-0.05, cex.axis=0.7)
box()
#Add line
lines(date, west, type="o", pch=11, lty=6, col="violet")
#labeling
title(main="War Data", col.main="green", font.main=3)
title(xlab="Date", col.lab=rgb(0,0.5,0))
title(ylab="War Readiness", col.lab=rgb(0,0.5,0))
#Legend
legend(-4000, y= 0.3, c("east","west"), cex=0.8, col=c("brown","violet"), pch=21:22, lty=1:2)
#PDF output
#dev.off()
west = morris_warmaking$West >  morris_warmaking$East
east = morris_warmaking$West <  morris_warmaking$East
equal = morris_warmaking$West ==  morris_warmaking$East

# Determine who was "better". I squeezed my brain for a better and fancy solution, but hours of thinking were pointless, so here is the simply not beautyful solution, I'm sorry. 

x = data.frame("date" = morris_warmaking$Date, "West_better" = west, "East_better" = east,  "Equal" = equal)
x


```

# Maximum Settlement Sizes 

> Fix the dataset so that it confirms to the principles of tidy data. What should be corrected?

* Get the whole thing in a table format with RegEx: `^(\W?\d+)\W(\S+)\s+([^\d]+)([^\s]+)\s+(.+)`
* The comma from the size column is interprated as decimal, so it has to be changed to . or nothing. 
* kill the author because he wrote "perhaps 500" in the western table for Mareybet and didn't asign a value for the points. In this case I just added the value NA to the table by myself and wrote the "perhaps" manually to the size, because I didn't want to change my regex just for a single entry
* One problem could be the 4 date of the western table because there the author starts with 8000 BCE. I would just ignore this and hope that the plot will to the right thing here and just start the eastern values at 4000 BCE 
* Commas are not supported in plot, therefore we have to change it and delete them
* Sometime there are inconsistencies like in the settlements column one time the author wrote ´London and Constantinople´ 

>The datasets for the East and the West are separate. What would be your strategies to combine them?

There are as always more ways to do this. I would clearly add another column with the header alignment (?!) to it and fill the cells with the values eastern/western, so I can you them in a single csv file. But out of convinience I just would import the files separetly for the visualisation. If we go really fancy and we don't have the alignment of the Settlements, I'm sure we could reference the Settlementnames with a gazetteer to get their geographical position and meta information. 


```{r} 
eastern <- read.csv("C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/settlements/morris_settlement_eastern.csv", sep=";")
western <- read.csv("C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/03/settlements/morris_settlement_western.csv", sep=";")
dates = western$Date
west_set = western$Settlement
west_size = western$Size
west_point = western$Points

east_set = eastern$Settlement
east_size = eastern$Size
east_point = eastern$Points

# Plot for the Population Size
plot(dates, east_size, type="o", col="brown", axes=FALSE, ann=FALSE, ylim=c(0,1000000))
axis(1, at=dates, lab=dates)
axis(2, at=east_size, las=1, tck=-0.05, cex.axis=0.7)
box()
## Add line
lines(dates, west_size, type="o", pch=11, col="violet")
## Labeling
title(main="Population Size", col.main="black", font.main=3)
title(xlab="Date", col.lab=rgb(0,0.5,0))
title(ylab="Size", col.lab=rgb(0,0.5,0))
## Legend
legend(-8000, y= 1000000, c("east","west"), cex=0.8, col=c("brown","violet"), pch=21:11, lty=1:2)


# Plot for the Warmaking Points
plot(dates, east_point, type="o", col="brown", axes=FALSE, ann=FALSE, ylim=c(0,10))
axis(1, at=dates, lab=dates)
axis(2, at=east_point, las=1, tck=-0.05, cex.axis=0.7)
box()
## Add line
lines(dates, west_point, type="o", pch=11, col="violet")
## Labeling
title(main="Warmaking Points", col.main="black", font.main=3)
title(xlab="Date", col.lab=rgb(0,0.5,0))
title(ylab="Points", col.lab=rgb(0,0.5,0))
## Legend
legend(-8000, y= 10, c("east","west"), cex=0.8, col=c("brown","violet"), pch=21:11, lty=1:2)

# Who was better?
west = west_point >  east_point
east = west_point <  east_point
equal = west_point == east_point


better_whole = data.frame("dates_lead" = dates, "West_better" = west, "East_better" = east,  "Equal" = equal, "Western_Cities" = west_set, "Eastern_Cities" = east_set)

# I think this is not a nice solution but I cannot make a valid if statement in a for each with data frames... this is frustrating... 

west_lead = subset(data.frame("dates_lead" = dates, "West_better" = west,    "Western_Cities" = west_set),  West_better == TRUE)
west_lead
east_lead = subset(data.frame("dates_lead" = dates, "East_better" = east,    "Eastern_Cities" = east_set),  East_better == TRUE)
east_lead

library("plyr")

total_lead = rbind.fill(west_lead, east_lead)
attach(total_lead)
newdata = total_lead[order(dates_lead),]
detach(total_lead)
newdata
```