---
title: "bosker"
output:
  html_document:
  df_print: paged
  pdf_document: default
  word_document: default
author: Bernhard Koschicek
---
# Bosker Dataset 

## Loading of the Dataset
```{r}
library(foreign)
bagdad_london_finalRestat <- read.dta("C:/Users/bkoschicek/Dropbox/Geschichte Studium/3_Sem/R-repo/04/Bosker_Data/Bosker_Data/bagdad-london--finalRestat.dta")
bagdad_london_finalRestat
#summary(bagdad_london_finalRestat$year)
#subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year==1800)
```

## What is the chronological extent of this data?

Chronological Begin:
```{r}
min(bagdad_london_finalRestat$year)
```

Chronological End:
```{r}
max(bagdad_london_finalRestat$year)
```

Summary: (Just for fun)
```{r}
summary(bagdad_london_finalRestat$year)
```

## What periods can it be divided into? How can we do that?
Well there are many humanists ways to seperate the data. Here it will just be divided in Medieval and Early Modern Era. We/I decided that it will be the best to split it at 1500, and 1500 is still medieval. 

```{r}
MedievalPeriod <- subset(bagdad_london_finalRestat,  bagdad_london_finalRestat$year <= 1500)
EarlyModernPeriod <- subset(bagdad_london_finalRestat,  bagdad_london_finalRestat$year > 1500)
head(MedievalPeriod)
tail(MedievalPeriod)
head(EarlyModernPeriod)
tail(EarlyModernPeriod)

```

## Can you generate a cumulative graph of population over time, divided into these periods? 
Here we see two graphs of the cumulative growth of the populations over time.
```{r}
# Create Data Frame for Medieval and Early Medieval Period
data.frame.Medieval <- data.frame("year" = MedievalPeriod$year, "country" = MedievalPeriod$country, "pop" = MedievalPeriod$total_pop_country)
data.frame.EarlyModern <- data.frame("year" = EarlyModernPeriod$year, "country" = EarlyModernPeriod$country, "pop" = EarlyModernPeriod$total_pop_country)

# Make them unique becuase the total population is states in every city, so just need one country per time. 
uniqueMedieval = unique(data.frame.Medieval[c("year", "country", "pop")])
aggMedieval = aggregate(uniqueMedieval$pop, by=list(Year=uniqueMedieval$year), FUN=sum)

uniqueEarlyModern = unique(data.frame.EarlyModern[c("year", "country", "pop")])
aggEarlyModern = aggregate(uniqueEarlyModern$pop, by=list(Year=uniqueEarlyModern$year), FUN=sum)

#Plot it!

plot(x = aggMedieval$Year, y = aggMedieval$x, type = "l", col="red", ylab = "Medieval Population", xlab = "Time", main = "Medival Population over Time")
par(new = TRUE)

legend("top", legend=c("Medival"),col=c("red"), lty = 1:2, cex=0.8)
```
```{r}
# Plot for Early Modern Population
plot(x = aggEarlyModern$Year, y = aggEarlyModern$x, type = "l", col="green", xlab = "Early Modern Population", ylab = "Time", lty = 2, main = "Early Modern population over Time")
par(new = TRUE)
legend("top", legend=c("Early Modern"),col=c("green"), lty = 1:2, cex=0.8)
```

```{r}
library(ggplot2)
ggplot(data = aggEarlyModern, aes(x = aggEarlyModern$Category, y = aggEarlyModern$x)) +
  geom_line(aes(y = aggEarlyModern$x, x= aggEarlyModern$Category), size=1, color ="green") +
  xlab("Year") + ylab("Population") +
  ggtitle("Early Modern Population") +
 	scale_x_continuous(breaks =round(seq(min(range_years_early_modern), max(range_years_early_modern), by = 100),1)) +
  scale_y_continuous(breaks = round(seq(min(aggEarlyModern$x), max(aggEarlyModern$x), by = 10000),1))
```
```{r}
library(ggplot2)
ggplot(data = aggMedieval, aes(x = aggMedieval$Category, y = aggMedieval$x)) +
  geom_line(aes(y = aggMedieval$x, x= aggMedieval$Category), size=1, color ="steelblue") +
  xlab("Year") + ylab("Population") +
  ggtitle("Medieval Population") +
  scale_x_continuous(breaks =round(seq(min(range_years_medival), max(range_years_medival), by = 100),1)) +
  scale_y_continuous(breaks = round(seq(min(aggMedieval$x), max(aggMedieval$x), by = 10000),1))
```
The code gives both graphs in one plot:
```{r}
medieval_population <- Medieval.bagdad.sub$citypop_le5
range_population_medieval <- range(Medieval.bagdad.sub$citypop_le5)
range_population_early_modern <- range(EarlyModern.bagdad.sub$citypop_le5)
range_years_early_modern <- range(EarlyModern.bagdad.sub$year)
years_medival_early_modern <- unique(bagdad_london_finalRestat$year)

library(ggplot2)
ggplot() +
  geom_line(data = aggMedieval, aes(y = aggMedieval$x, x= aggMedieval$Category), size=1, color ="steelblue") +
  geom_line(data= aggEarlyModern, aes(y = aggEarlyModern$x, x= aggEarlyModern$Category), size= 1, color ="green") +
  xlab("Year") + ylab("Population") +
  ggtitle("Population over Time") +
  scale_x_continuous(breaks =round(seq(min(years_medival_early_modern), max(years_medival_early_modern), by = 100),1)) +
	scale_y_continuous(breaks = round(seq(min(aggMedieval$x), max(aggEarlyModern$x), by =10000),1)) +
  scale_color_discrete(name="Period",  labels = c("Early Modern Population", "Medival Population"))
```




# North Africa and Europe?
## Can you construct comparative graphs of population in North Africa and Europe (similar to what you did with the Morris dataset). Here you will need to sum up population!
First split the data frame in a African and European Data Frame.

```{r}
# Create Vector for Europe
Europevector <- c("Austria", "Ireland", "Belgium", "Czech rep.", "Denmark", "Finland", "France", "Germany", "Hungary", "Italy", "Netherlands", "Norway", "Poland", "Portugal", "Slovakia", "Spain", "Sweden", "UK", "Switzerland", "Greece", "Turkey", "Yugoslavia", "Albania", "Bulgaria", "Rumenia", "Luxemburg", "Malta")

# Create Vector for Afrika 
NorthAfricavector <- c("Morocco", "Tunesia", "Lybia", "Egypt", "Syria")

# Create the Europe Data Frame
Europe.bagdad <- bagdad_london_finalRestat$country %in% Europevector

europe <- bagdad_london_finalRestat[bagdad_london_finalRestat$country %in% Europevector, ]

# Create the North Afrika Data Frame

NorthAfrica.bagdad <- bagdad_london_finalRestat$country %in% NorthAfricavector

northAfrica<- bagdad_london_finalRestat[bagdad_london_finalRestat$country %in% NorthAfricavector, ]
```

Now we can aggregate the whole thing to collapse the data and plot it in a beautyful plot. 

```{r}
# Aggregate European Data
europe.frame = data.frame("year" = europe$year, "country" = europe$country, "pop" = europe$total_pop_country)
uniqueEurope = unique(europe.frame[c("year", "country", "pop")])
aggEurope = aggregate(uniqueEurope$pop, by=list(Year=uniqueEurope$year), FUN=sum)
aggEurope

# Aggregate African Data
africa.frame = data.frame("year" = northAfrica$year, "country" = northAfrica$country, "pop" = northAfrica$total_pop_country)
uniqueAfrica = unique(africa.frame[c("year", "country", "pop")])
aggAfrica = aggregate(uniqueAfrica$pop, by=list(Year=uniqueAfrica$year), FUN=sum)
aggAfrica

# Plot the whole thing 

plot(x = aggAfrica$Year, y = aggAfrica$x, type = "l", col="red", ylab = "Population per country", xlab = "Time", main = "Comparative graph of Europe's and North Africa's population")
par(new = TRUE)

plot(x = aggEurope$Year, y = aggEurope$x, type = "l", col="green", axes = FALSE, xlab = " ", ylab = " ", lty = 2)

legend("top", legend=c("North Africa", "Europe"),col=c("red", "green"), lty = 1:2, cex=0.8)

```


# Habsburg Empire 1500 - 1800

```{r}

# Habsburg at the 15 cent. Just country selection, but the way you do it is the same 
habsburg_15_century <- c("Austria", "Spain", "France", "Italy","Belgium", "Netherlands", "Yugoslavia")
habsburg_frame_15 <- subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1500)
habsburg_frame_15 <- habsburg_frame_15[(habsburg_frame_15$country %in% habsburg_15_century), ]
habsburg_frame_15 %>% count(habsburg_frame_15$year)

# Habsburg at the 16 cent. Just country selection, but the way you do it is the same 
habsburg_16_century <- c("Austria", "Hungary", "Italy", "Slovakia", "Czech rep.", "Rumenia", "Serbia", "Croatia", "Slovenia", "Galicia", "Ukraine", "Poland", "Luxemburg", "Belgium", "Germany", "France", "Spain", "Portugal", "Netherlands")
habsburg_frame_16 <- subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1600)
habsburg_frame_16 <- habsburg_frame_16[(habsburg_frame_16$country %in% habsburg_16_century), ]
habsburg_frame_16 %>% count(habsburg_frame_16$year)


```

# islamdom and Christiandom

# ```{r}
# #The ugliest way to determine the biggest muslim cities in Europe between 1000 and 1500.
# IslamEurop.frame.1000 <- IslamEurop.frame[IslamEurop.frame$year >= 1000, ]
# IslamEurop.frame.1500 <- IslamEurop.frame.1000[IslamEurop.frame.1000$year <= 1500, ]
# IslamEurope.fram.1500.Islam <- IslamEurop.frame.1500[IslamEurop.frame.1500$muslim == 1, ]
# IslamEurope.fram.1500.Islam.Pop <- IslamEurope.fram.1500.Islam[IslamEurope.fram.1500.Islam$Citypop > 20, ]
# IslamEurope.fram.1500.Islam.Pop.Oops <- IslamEurope.fram.1500.Islam.Pop[IslamEurope.fram.1500.Islam.Pop$country != "Turkey", ]
# 
# unique(IslamEurope.fram.1500.Islam.Pop.Oops$city)
# 
# #and a clunky way to do the biggest cities both before and after 1500, our period split
# 
# Islam.Beta <- data.frame("year" = bagdad_london_finalRestat$year, "city" = bagdad_london_finalRestat$city, "CityPop" = bagdad_london_finalRestat$citypop_le10, "muslim" = bagdad_london_finalRestat$muslim)
# 
# Islam.Frame <- Islam.Beta[Islam.Beta$muslim == 1, ]
# 
# Islam.Frame.Pre1500 <- Islam.Frame[Islam.Frame$year < 1499, ]
# Islame.Frame.Post1500 <- Islam.Frame[Islam.Frame$year >= 1500, ]
# 
# Islam.Frame.Pre1500.Pop <- Islam.Frame.Pre1500[Islam.Frame.Pre1500$CityPop > 20, ]
# Islam.Frame.Post1500.Pop <- Islam.Frame.Pre1500[Islam.Frame.Pre1500$CityPop > 20, ]
# 
# unique(Islam.Frame.Pre1500.Pop$city)
# 
# unique(Islam.Frame.Post1500.Pop$city)

```


```{r}
#Largest cities of Islamdom for each reported period
head(bagdad_london_finalRestat)
acht<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 800)
neun<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 900)
t<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1000)
teins<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1100)
tzwei<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1200)
tdrei<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1300)
tvier<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1400)
tfunf<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1500)
tsechs<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1600)
tsieben<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1700)
tacht<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$year == 1800)
#brauch ich auch nicht: Muslimcit<-subset(bagdad_london_finalRestat, bagdad_london_finalRestat$muslim == 1)
achtmuslcit<-subset(acht, acht$muslim==1)
max(achtmuslcit$citypop_le10)
neunmuslcit<-subset(neun, neun$muslim==1)
max(neunmuslcit$citypop_le10)
tmuslcit<-subset(t, t$muslim==1)
max(tmuslcit$citypop_le10)
teinsmuslcit<-subset(teins, teins$muslim==1)
max(teinsmuslcit$citypop_le10)
tzweimuslcit<-subset(tzwei, tzwei$muslim==1)
max(tzweimuslcit$citypop_le10)
tdreimuslcit<-subset(tdrei, tdrei$muslim==1)
max(tdreimuslcit$citypop_le10)
tviermuslcit<-subset(tvier, tvier$muslim==1)
max(tviermuslcit$citypop_le10)
tfunfmuslcit<-subset(tfunf, tfunf$muslim==1)
max(tfunfmuslcit$citypop_le10)
tsechsmuslcit<-subset(tsechs, tsechs$muslim==1)
max(tsechsmuslcit$citypop_le10)
tsiebenmuslcit<-subset(tsieben, tsieben$muslim==1)
max(tsiebenmuslcit$citypop_le10)
tachtmuslcit<-subset(tacht, tacht$muslim==1)
max(tachtmuslcit$citypop_le10)

library(dplyr)
cityacht<-filter(achtmuslcit, achtmuslcit$citypop_le10>=350)
cityacht
cityneun<-filter(neunmuslcit, neunmuslcit$citypop_le10>=450)
cityneun
citytausend<-filter(tmuslcit, tmuslcit$citypop_le10>=300)
citytausend
citytausendein<-filter(teinsmuslcit, teinsmuslcit$citypop_le10>=250)
citytausendein
citytausendzwei<-filter(tzweimuslcit, tzweimuslcit$citypop_le10>=200)
citytausendzwei
citytausenddrei<-filter(tdreimuslcit, tdreimuslcit$citypop_le10>=220)
citytausenddrei
citytausendvier<-filter(tviermuslcit, tviermuslcit$citypop_le10>=250)
citytausendfunf<-filter(tfunfmuslcit, tfunfmuslcit$citypop_le10>=280)
citytausendsechs<-filter(tsechsmuslcit, tsechsmuslcit$citypop_le10>=700)
citytausendsieben<-filter(tsiebenmuslcit, tsiebenmuslcit$citypop_le10>=700)
citytausendacht<-filter(tachtmuslcit, tachtmuslcit$citypop_le10>=500)
citytausendacht
```

